### 2011-11-07 inst/doc/article.Rnw
Article

{\color{blue} \tt \bfseries $^{1}$ Groningen Bioinformatics Centre - University of Groningen \\
$^{2}$ Genomics Coordination Centre \& University Medical Centre Groningen \\
version: \# 0.0.1\\[24pt]
received on XX XXX 201X\\
accepted on XX XXX 201X\\
}

@@

\textbf{\sffamily Abstract}: In this paper we present Quanitative Correlation Locus 
(QCL) mapping, a novel approach to detect genetic regulation of phenotypes in 
recombinant inbred line populations (RIL) or other experimental crosses. QCL mapping 
allows association of phenotype correlation differences to genetic markers, in almost 
the same way as QTL mapping associates genetic markers with differences in expression.
We use QCL mapping to reconstruct underlying genetic architecture, and provide insight 
into regulation of (gene) expression. QCL information in combination with classical QTL 
information leads to a more detailed insight into the genetic architecture underlying 
complex traits. 
In two example we reconstruct part of the genetic architecture in S. Cerevisiae grown 
under ethanol conditions. The other example shows the reconstruction of the chain 
elongation pathway in A. Thaliana. In both examples we map the observe differences 
in correlation to genetic markers, and show the biological relevance of QCL peeks and 
hotspots in regulatory network reconstruction.

@@

\noindent Corresponding Author: \verb|Danny.Arends@Gmail.com|
\noindent Associate Editor: XXXX XXXX
\newpage

@@

S. Cerevisiae dataset from the gene expression omnibus (GEO) (GDS1115 / GDS1116). This 
dataset contains expression profiles of parental strains and haploid progenies from a 
cross of strain BY4716 and the wild wine strain RM11-1a. Data was collected for 4482 
probes. Genotype data for the 109 progenies was available at 282 genetic markers 
obtained by contacting the authors of the original paper. Expression of probes was 
measured under growth conditions containing 1% ethanol. Additional details about the 
experimental settings are available in the original paper (Brem et al. 2005).

@@

Global overview of QCL scan after using a full scan for QCL
<<echo=FALSE>>=
  qcl_result
@ 
Single results from a QCL mapping, shows the pattern at the highest threshold.
<<echo=FALSE>>=
  qcl_result[[2]]
@ 
Phenotype x Marker image plot, this is 'comparable' to a QTL scan on mutliple traits.

@@

Phenotype x Phenotype image plot, the new network information obtained by QCL mapping, 
which can be used to reconstruct the underlying genetic architecture.
\begin{figure}[ht]
\centering
<<p2pmatrix, fig=TRUE, echo=FALSE, height=3.5>>=
par(mar=c(3,8,1.1,1.1))
image(qcl_result, 0.5, against="phenotypes")
@ 
\caption{Phenotypes x Phenotypes matrix (A. Thaliana).\label{fig:p2pmatrix}}
\end{figure}

@@

Microarray experiments have allowed the identification of genes differentially 
expressed between various conditions, or groups of genes coexpressed across 
samples. Many genetical genomics screens have been carried out to map differential 
expression to genetic variation. Screening the genome for genotype combinations 
for which there is differential expression has been a successful strategy when 
mapping Mendelian traits. Results on complex traits have been however less then 
satisfactory with only small amount of heritability explained by individuals loci. 
QCL mapping however 'borrows' information from each phenotype in the analysis set, 
and enables easy extraction of phenotype marker phenotype networks. furthermore QCL 
seem to 'transfer' between phenotypes in a more stabile fashion which gives greater 
detail and confidence when reconstructing genetic architecture.

@@

Langfelder P, Horvath S (2008) WGCNA: an R package for weighted correlation network analysis. BMC Bioinformatics 2008, 9:559

>>>>

{\color{blue} \tt \bfseries $^{1}$ Groningen Bioinformatics Centre - University of Groningen\\
$^{2}$ Genomics Coordination Centre - University Medical Centre Groningen\\
version: \# 0.0.1\\[24pt]
received on XX XXX 201X\\
accepted on XX XXX 201X\\
}

@@

\textbf{\sffamily Abstract}: In this paper we present Quanitative Correlation Locus 
(QCL) mapping, a novel approach to detect genetic regulation of phenotypes in 
recombinant inbred line populations (RIL) or other experimental crosses. QCL mapping 
allows association of phenotype correlation differences to genetic markers, in almost 
the same way as QTL mapping associates genetic markers with differences in expression.
We use QCL mapping to reconstruct underlying genetic architecture, and provide insight 
into regulation of (gene) expression. QCL information in combination with classical QTL 
information leads to a more detailed insight into the genetic architecture underlying 
complex traits. 
In two example we reconstruct part of the genetic architecture in S. Cerevisiae, another 
example shows the reconstruction of the chain elongation pathway in A. Thaliana. In both 
examples we are able to map the observe differences in correlation to genetic markers, 
and show the biological relevance of QCL peeks and hotspots in regulatory network 
reconstruction.

@@

\noindent Corresponding Author: Danny Arends \verb|Danny.Arends@Gmail.com|
\noindent Associate Editor: XXXX XXXX
\newpage

@@

S. Cerevisiae dataset from the gene expression omnibus (GEO) (GDS1115 / GDS1116). This 
dataset contains expression profiles of parental strains and haploid progenies from a 
cross of strain BY4716 and the wild wine strain RM11-1a. Data was collected for 4482 
probes. Genotype data for the 109 progenies was available at 282 genetic markers 
obtained by contacting the authors of the original paper. Expression of probes was 
measured under growth conditions containing 1\% ethanol. Additional details about the 
experimental settings are available in the original paper (Brem et al. 2005).

@@

Global overview of QCL scan after using a full scan for QCL
<<echo=FALSE>>=
  qcl_result
@ 
Single results from a QCL mapping, shows the pattern at the highest threshold.
<<echo=FALSE>>=
  qcl_result[[2]]
@

\emph{\sffamily The Phenotypes x Marker relations}

Phenotype x Marker image plot, this is 'comparable' to a QTL scan on mutliple traits.

@@

\emph{\sffamily The Phenotypes x Phenotypes relations}

Phenotype x Phenotype image plot, additional information obtained by QCL mapping, 
which can be used to reconstruct the underlying genetic architecture.
\begin{figure}[ht]
\centering
<<p2pmatrix, fig=TRUE, echo=FALSE, height=3.5>>=
par(mar=c(3,8,1.1,1.1))
image(qcl_result, 0.5, against="phenotypes")
@ 
\caption{Phenotypes x Phenotypes matrix (A. Thaliana).\label{fig:p2pmatrix}}
\end{figure}

@@

Microarray experiments have allowed the identification of genes differentially 
expressed between various conditions, or groups of genes coexpressed across 
samples.

Many genetical genomics screens have been carried out to map differential 
expression to genetic variation. Screening the genome for genotype combinations 
for which there is differential expression has been a successful strategy when 
mapping Mendelian traits. However results on complex traits have been less 
then satisfactory with only small amounts of heritability explained by 
individuals loci. 
QCL mapping however 'borrows' information from each phenotype in the analysis set, 
and enables extraction of phenotype to marker to phenotype networks. 
Furthermore QCLs seem to 'transfer' from phenotype to phenotypes in a stabile 
fashion, which gives greater detail and confidence when reconstructing genetic 
architecture. 
It can provides better insight in the missing of heritability of variation, we 
hypothesize that changes in variation are affecting (by loss of) correlation in 
the normally correlated phenotypes leading to an amplification of this variation, 
by distrupting the network environment of which endo-phenotypes are part of.

@@

Langfelder P, Horvath S (2008) WGCNA: an R package for weighted correlation network 
analysis. \emph{BMC Bioinformatics}, 9:559
